---
title: "R Notebook"
---

## Scenario


The data here are used by an app that enables clinicians to monitor their patients' medication prescriptions.

There is a __"narrative"__ field that contains formatted text that is ready for display in a browser.
There are some concerns of problematic entries in the narrative field that are causing confusion for the clinicians - it should match the __"coding_display"__ field, but dos not in all cases.


**************************************************


## Brief

Being somewhat knowledgeable about the FHIR API, and having some decent R skills, it has fallen to you to investigate the issue.  
You will have to  

- identify the problematic entries
- fix any issues
- identify the patients by name, to allow the clinicians to contact any patients they think might be adversely affected.

You have access to the following two data frames (you will have to adjust the path as necessary).

```{r}
library(dplyr)
library(purrr)
library(stringr)
library(httr2)

# Might need to adjust path depending on where you're running it from
load('week-04-tutorial.RData')
# Loads two dataframes
# - smart_patients
# - smart_requests (i.e med requests)
```

The __narrative__ field is in the `text_div` column in the `smart_requests` dataframe.

Example of narrative field  
`<div html stuff>  MEDICATION DETAILS [rxnorm 123456]  </div>`  

The text contained in the narrative column is an __HTML formatted string__ that is displayed by the app, to the clinician.


The narrative text is automatically populated with the information contained in the `coding_display` column.
Therefore, the text in the `text_div` and `coding_display` columns should match.


**************************************************


## Primer on html tags

HTML formatting consists of paired tags - an opening tag, and a closing tag.

Tags themselves are indicated by opening and closing angle brackets `<tag>` (opening) or `</tag>` (closing, indicated by the `/`).
The tag found in the narrative column is called `<div>`.

All content found __between the opening and closing tags__ is the patients' medication information, e.g.  
`<div html stuff>  MEDICATION DETAILS [rxnorm 123456]  </div>`  
The rxnorm code should also be excluded, so in the above example only "MEDICATION DETAILS" should be extracted.


**************************************************


## The task

The steps required to complete the task are as follows

1. Text consistency validation
    1. __Write a test__ that will pass when supplied two text vectors that match - i.e. the narrative, and the display columns
    1. __Process the narrative div__ to match display field by removing the HTML tags.
    1. The `[rxnorm 123456]` component is not needed and should also be __removed__.
    1. Identify the records with non-matching entries in the narrative field.

1. Identify any affected patients.
    1. __Write a test__ that ensures patient IDs conform to the pattern `"smart-..."`.
    1. __Extract or replace__ any suspected typos.
    1. __Join__ with the __patient info data__ to get the names and adresses of the affected patients.


## Write test

Write a test that will compare two string vectors and return `TRUE` if they all match and `FALSE` otherwise.
Whether you want to use `testthat` or not is up to you.

```{r}

```


## Extract medication text from div

Use any of the string processing methods discussed on the course this week as required to __extract medication text from div__.
Place the processed result in a new column for easy inspection.

```{r}
smart_requests |> 
  mutate(______ = ______,
         ...... = ??????)
```


#### Apply test

Use your test to confirm all records now match.
You can copy it, and supply the necessary columns as inputs in any way you like.
You may have to examine the results and iterate.

```{r}

```


## Filter for only the affected patients

```{r}
smart_requests |> 
  filter(______)
```

## Join with the patient data

```{r}
smart_patients |> 
  ***_join(_______)
```

