---
title: "Patient report"
format: html
---

```{r}
#| label: setup
#| include: false

knitr:::opts_chunk$set(include = FALSE)

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(httr2)

```

## Create API request

### Set API endpoints

```{r}
#| label: set-api-parameters

base_url <- 'https://lforms-fhir.nlm.nih.gov/baseR4'
resource <- 'Patient'
patient_id <- 'pat-106'
```

You can a different patient ID.

- pat-106
- pat-2338
- pat-2374
- pat-3688
- smart-1032702
- smart-1081332
- smart-1098667
- smart-1134281


### Patient request

build up the httr2 request object.
You will need to insert the API parameters above into the correct functions below.

```{r}
#| label: build-GET-request
#| eval: true
#| include: false

# ---- Ex 1.1 ----
# Fill in the function arguments. Use the help files, and discuss.

req_patient <- request() |>
  req_url_path_append() |>
  req_headers(Accept = 'application/fhir+json')
# Discuss the meaning of the final line above


# ---- Ex 1.2 ----
# Insert your request into the `dry_run()` function, and run it

req_dry_run()

# Examine the printed information - see if you can identify all the components of your request.


```



### Send the query, and explore the result

Use `req_perform()` to send your query from above to the servers.
Remember to assign the return value to a variable.
If the code runs without error, examine the response object, and locate the __status code__ - or access it directly using `$` and the name `status_code` - it should be __200__.

```{r}
#| label: run-query

# ---- Ex 1.3 ----
# Run query here see `req_perform()` and check response status.
# Perform the rquest and assign the value to a variable called `resp_patient`.


```

### Examine response body

Because our GET request header specified that we were ready to 'Accept' a JSON encoded response, the server should respond with a body in JSON format.
However, if we examine the response we see what looks like a long sequence of hexadecimal numbers
This is our data in raw byte format - this is done mainly for efficiency reasons.
When converted from this raw byte format (often encoded in UTF-8), it will become a readable JSON string.
We accomplish the conversion using `resp_body_json()`.

```{r}
#| label: convert-response-to-json

# ---- Ex 1.4 ----
# To convert your response object from above
# use the function resp_body_json() here




# Remember to store the result in a variable

```


## Extracting specific details

We will extract all the details we need for our report at the start of our document.

We can then refer to the variables as needed throughout.

```{r}
#| label: extract-name-info-1

# From your "converted data" object (last chunk)...

# Ex 2.0
# Find the patient's surname (i.e. their "family" name)
# Assign it to a variable called `family_name`


```

Use the example above as a template to explore other features of the  data.
You could also consider writing a "getter" function for extracting each individual component of the data.
For example, you could write a function called `get_family()` that simply accessed the family name.
You could then store that function in a script you sourced for every FHIR patient you analysed.

```{r}
#| label: extract-name-info-2

# Ex 2.1
# Find their first name (i.e. their "given" name)
# Assign it to a variable called `first_name`




# Ex 2.2
# Paste their names together and assign to a variable called `patient_name`




```


```{r}
#| label: compute-patient-age

# Ex 3.1
# Find their birth date
# Assign it to a variable called `birth_date`




# Ex 3.2
# Use the appropriate R functions to calculate their age
# I.e. subtract from the current year
# Note that although this is correct for calculating their age,
# the patient was born in the future, so this is actually "years til birth"




```


```{r}
#| label: patient-address

# Ex 4.1
## The patient's address is stored line by line, but also in an element called `$text$div
# Extract it from the `div` element, and save it in a variable called `patient_address



```

**************************************************

<!-- ############################################# -->
<!-- REPORT STARTS HERE -->
<!-- Render the document, and scroll to the bottom -->
<!-- ############################################# -->

## Patient Report

__Name:__ `r patient_name`  
__Age:__ `r patient_age`  

__Address:__ `r patient_address`


## Address from `$div` element


```{r}
#| label: pre-formatted-text
#| include: true

# The `HTML()` function will mark the text as HTML for RMarkdown

htmltools::HTML(resp_list$text$div)

```


## Some medical info

Follow the above and insert some simple information from one of the other FHIR resources, `Observation`, `Encounter`, or `MedicationRequest`.
This will require a new API request.

This is where things get tricky without an in depth knowledge of FHIR.
Not all of the following have information in all of the resources (I don't think?).

- pat-106
- pat-2338
- pat-2374
- pat-3688
- smart-1032702
- smart-1081332
- smart-1098667
- smart-1134281

__Patient resource__

They all will work with as this resource type using `...path_append('Patient', id)`

__Encounter resource__

They can all be found with using `req_url_query(subject = id)`

__Observation resource__

They can all be found with using `req_url_query(subject = id)`

__MedicationRequest resource__

Only the IDs beginning with `smart` can be ound using `req_url_query(patient = id)`


## For bonus points 

...see if you can create a table of several observations, encounters, or medication requests.

A nice "cheat" method for doing this is to include `simplifyVector = TRUE` in the call to `resp_json_body()`
