---
title: "Tutorial 02 Solutions Document: Intermediate R Programming"
output: html_document
date: "`r Sys.Date()`"
---

In this practice document we will be applying some of the core concepts and principles introduced this week into some hands on exercises. These course concepts include **reprex**, **debugging** and using the `{purrr}` package. 

We will need the following packages: 

```{r libraries, message=FALSE}
library(purrr)
library(tibble)
library(dplyr)
library(reprex)
```

## Exercise 1

We have a data object called `moons` which has the diameter of the moons of some planets in our solar system. 

How can you get the length of each vector in the list `moons`? This will tell us how many moons the planets Earth, Mars, and Neptune have. 

The output of this should be an integer vector. 

```{r ex-1-data}
# values represent the diameter of each moon 
moons <- list(
  earth = 3474.8,
  mars = c(22.53, 12.4),
  neptune = c(150, 60, 42, 2706.8, 62, 340, 18,
              38, 420, 82, 44, 194, 174.8, 66)
)
```

### Solution 

We cannot just do `length(moons)` as this returns the number of elements in the list (AKA the number of planets). We need to use `map` to iterate through the vectors in the list and get the length of each element. 

```{r ex-1-solution}
# does not give us the answer we are looking for
length(moons)

# we get 4 because there are 4 elements in our list!
# using map does give us what we are looking for
moons |>
  map(length) # outputs a list

# to get an integer vector we need to use map_int
moons |>
  map_int(length)
```

We could also use another function `lengths()` which is a bit of a game changer, it gives "the length of each element of a list or atomic vector as an integer or numeric vector" (check the documentation for more.)

```{r ex-1-lengths}
#lengths(moons)
moons %>%
  lengths()
```

## An aside on data types and structures

We discussed briefly this week the difference between `NA` and `NULL` in Topic 1.2. If we wanted to add Venus to our `moons`list, which missing value is most appropriate and accurate? 

Notice that we created `moons` as a list. Can you think of a reason as to why this data may be better represented as a list than as a more traditional data frame or tibble? An answer will be provided after Exercise 3. 

## Exercise 2 

Using the same object `moons`, which planet has the smallest median moon radius? 

### Solution 

Mars! 

```{r ex-2-solution}
moons |>
  map_dbl(median)
```

## Exercise 3 

At the moment, the values in our list `moon` are in kilometers. Change them to miles using an anonymous function. 

*Hint* To convert km to miles, multiple by 0.62 

### Solution 

This is an ideal use case for an anonymous function, as after we do the conversion, it is unlikely we will need the function again. 

```{r ex-3-solution}
moons_mi <- moons |>
  map(\(value) value * 0.62)

moons_mi
```

## An answer on data types and structures 

We are using `map()` in these exercises to *loop* over a list, and apply a *vectorized* function. This speaks to the change in mind-set required to think in both a functional **and** vectorized way. 

Venus does not have any moons. `NA` returns a length of 1, so the correct missing value here is `NULL`. `NA` is more appropriate when we do not have the data for something, but it does theoretically exist. Here, no moon exists it is not that we are missing the value of the diameter. 

However, I have not added Neptune into our working `moons` object because working with objects of length 0 can cause troubles we do not want to deal with today. If you are interested in exploring this more, add `venus = NULL` in exercise 1 where we create the `moons` list and work through exercises 1 to 3 and see what changes. 

```{r missing-data}
length(NA) #> 1

length(NULL) #> 0
```

Tibbles must have columns (which are vectors) of the same length. Lists on the other hand have no such restriction around the length of elements they contain. Neptune has 14 moons, which would mean a lot of `NA`s for Earth and Mars which only have 1 and 2 moons, respectively. 

Lets explore this... 

## Exercise 4 

Run the following code and read the error message. We already know what is wrong here, but try and understand how the error message is communicating this to you. 

```{r ex-4-error}
moons_df <- tibble(
  venus = NULL,
  earth = 3474.8,
  mars = c(22.53, 12.4),
  neptune = c(150, 60, 42, 2706.8, 62, 340, 18,
              38, 420, 82, 44, 194, 174.8, 66)
)
```

### Solution 

Straight away we see at the to that `"! Tibble columns must have compatible sizes."`, which is telling us the issue we already know about - our columns are not all 14 rows/elements long! 

An interesting note here is that `"ℹ Only values of size one are recycled."`

**Recycling** describes the concept of repeating elements of one vector to match the size of another. There are two rules that underlie the `{tidyverse}` recycling rules:

1. Vectors of size 1 will be recycled to the size of any other vector
2. Otherwise, all vectors must have the same size

So if we run the code from above with only Earth and Neptune, we actually do not get an error! However... the result may not be what you were anticipating. The **key point** here is that it is important to always check your objects after applying some function to them, to ensure the result is what you expected. The most difficult errors to catch are the silent ones that execute, just not doing what you wanted. 

```{r ex-4-solution}
moons_en_df <- tibble(
  venus = NULL,
  earth = 3474.8,
  #mars = c(22.53, 12.4),
  neptune = c(150, 60, 42, 2706.8, 62, 340, 18,
              38, 420, 82, 44, 194, 174.8, 66)
)

dplyr::glimpse(moons_en_df)
```

## Exercise 5

The following code uses `map` nested inside another `map` to apply a function to every element of a nested list. Why does it fail, and what do you need to do to make it work? 

Use debugging tools such as reading the error message, traceback, and/or the interactive debugger to help you answer these questions. 

```{r ex-5-data}
ex_list <- list(
  list(1, c(3, 9)),
  list(c(3, 6),  7, c(4, 7, 6))
)

str(ex_list)
```

```{r ex-5-error}
triple <- function(x) x * 3

map(ex_list, map, .f = triple)
#> Error in `map()`:
#> ℹ In index: 1.
#> Caused by error in `.f()`:
#> ! unused argument (function (.x, .f, ..., .progress = FALSE)
#> {
#>    map_("list", .x, .f, ..., .progress = .progress)
#> })
```

### Solution 

This function call fails, because `triple()` is specified as the `.f` argument and consequently belongs to the outer/1st `map()`. The unnamed argument `map` is treated as an argument of `triple()`, which causes the error.

There are at least 2 ways you can solve this:

```{r ex-5-solution}
## Option 1: do not name the argument
map(ex_list, map, triple)

## Option 2: Use anonymous functions!
map(ex_list, \(x) map(x, triple))
```

If you came up with another solution, share it on the discussion boards! 

## Exercise 6 

Look up the documentation for `purrr::map2()`. Read the help files to understand the arguments and what the function does. Perhaps even try out some of the examples at the bottom of the help file. 

Then apply this function to do the follow task: 

Imagine we have some data on the number of different fruit that you ate each day across a week. Create a new column in `fruits_num` called `total_fruit` which is the sum of each row in columns `apples` and `bananas`. 

**Hint** What function from `{dpylr}` do we use to "create, modify and delete columns"?  

**Hint 2** Remember, tibble columns are vectors

**Hint 3** Think about what output data structure you would like, is it a list? Or a vector of a specific data type? 

```{r ex-6-data}
fruits_num <- tibble(
  # create days of the week
  # if you are not sure what the arguments to ISOdate() are, look up the documentation! 
  days = weekdays(ISOdate(1, 1, 1:7)),
  apples = c(1, 2, 4, 6, 2, 1, 8),
  bananas = c(6, 5, 3, 2, 1, 9, 4)
)
```

### Solution 

Because tibble columns are vectors, so we can use `map2()` inside `mutate()` to alter tibble columns!

```{r ex-6-solution}
# Option 1 with base::sum()
fruits_num |>
  mutate(total_fruit = map2_dbl(.x = apples,
                                .y = bananas,
                                .f = sum))

# Option 2 with anonymous function
fruits_num |>
  mutate(total_fruit = map2_dbl(.x = apples,
                                .y = bananas,
                                .f = \(x, y) x + y))
```

If we used `map2` we would end up with a list-column for the sum, rather than a column which is a numeric vector: 

```{r ex-6-solution-2}
fruits_num |>
  mutate(total_fruit = map2(.x = apples,
                            .y = bananas,
                            .f = sum))
```

This is not necessarily incorrect, but is unlikely to be what you were wanting as output and complicates any handling of the tibble in downstream analysis. 

## Exercise 7

Create a reprex of your solution to Exercise 6 and post it on the dedicated discussion board in the week 2 folder. This is an exciting step, your (likely) first reprex!  

*Hint:* Use the `{reprex}` package to ensure your solution is a reprex. See week 2 Topic 3.5 in the course content. 

### Solutions 

 See `Tutorial-Week_2_EX_7.R` and solutions video 
